# ARD-first tables

## ARD-first Tables

Similar to functions that accept a data frame, the package exports functions with nearly identical APIs that accept an ARD.

::: columns
::: {.column width="50%"}

```r
tbl_summary()

tbl_continuous()

tbl_wide_summary()
```

:::
::: {.column width="50%"}

```r
tbl_ard_summary()

tbl_ard_continuous()

tbl_ard_wide_summary()
```

:::
:::

## ARD-first Tables

We can use the skills we learned earlier today to create ARDs for gtsummary tables.

```{r}
#| label: 'ard_stack'
#| output-location: "column"
library(cards)

ard <- ard_stack(
  data = trial, 
  ard_continuous(variables = age),
  ard_categorical(variables = grade),
  ard_dichotomous(variables = response),
  # add these for best-looking tables
  .attributes = TRUE, 
  .missing = TRUE 
)
ard
```

::: aside
The `.attributes=TRUE` call adds column attributes, like labels, to the ARD table, which leads to better defaults by displaying the column labels in the table.
:::

## ARD-first Tables

We can simply use the ARD from the previous slide, and pass it to `tbl_ard_summary()` for a summary table.

```{r}
#| label: 'tbl_ard_summary'
#| output-location: "column"

tbl_ard_summary(ard)
```


## ARD-first Tables

Now let's try a somewhat more complicated table.

```{r}
#| label: 'tbl_ard_summary2'
#| output-location: "slide"
#| code-line-numbers: "|2|4|7|17|18|19|21"
trial |> 
  labelled::set_variable_labels(age = "Age, years") |> 
  ard_stack( 
    .by = trt,
    ard_continuous(
      variables = age,
      fmt_fn = age ~ list(sd = 2)
    ),
    ard_categorical(variables = grade),
    ard_dichotomous(variables = response),
    # add these for best-looking tables
    .attributes = TRUE, 
    .missing = TRUE 
  ) |> 
  tbl_ard_summary(
    by = trt,
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{mean} ({sd})", "{min} - {max}"),
    missing = "no"
  ) |> 
  modify_caption("**Table 1. Subject Demographics**")
```

## ARD-first Table *Shells*

```{r}
#| label: 'tbl_ard_summary-mock'
#| output-location: "slide"
#| code-line-numbers: "15"
trial |> 
  labelled::set_variable_labels(age = "Age, years") |> 
  ard_stack( 
    .by = trt,
    ard_continuous(
      variables = age,
      fmt_fn = age ~ list(sd = 2)
    ),
    ard_categorical(variables = grade),
    ard_dichotomous(variables = response),
    # add these for best-looking tables
    .attributes = TRUE, 
    .missing = TRUE 
  ) |> 
  dplyr::mutate(fmt_fn = list(\(x) "xx.x")) |> 
  tbl_ard_summary(
    by = trt,
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{mean} ({sd})", "{min} - {max}"),
    missing = "no"
  ) |> 
  modify_header(all_stat_cols() ~ "**{level}**  \nN = xx")
```


## {gtsummary} Exercise 2

1. Navigate to Posit Cloud script `05-gtsummary_exercise2.R`.

2. Create the table outlined in the script.

3. Add the "completed" sticky note to your laptop when complete.

```{r}
#| echo: false
#| cache: false
countdown(minutes = 10)
```
