# {gtsummary} themes

## {gtsummary} theme basics

- A [theme]{.emphasis} is a set of customization preferences that can be easily set and reused. 

- Themes control [default settings for existing functions]{.emphasis}

- Themes control more [fine-grained customization]{.emphasis} not available via arguments or helper functions

- Easily use one of the [available themes]{.emphasis}, or [create your own]{.emphasis}

## {gtsummary} default theme

```{r}
#| output-location: "column"
reset_gtsummary_theme()

trial |> 
  tbl_summary(
    by = trt, 
    include = c(age, response)
  ) |>
  modify_caption(
    "Default Theme"
  )
```

## {gtsummary} theme_gtsummary_journal()

```{r}
#| output-location: "column"
reset_gtsummary_theme()
theme_gtsummary_journal(journal = "jama")

trial |> 
  tbl_summary(
    by = trt, 
    include = c(age, response)
  ) |>
  modify_caption(
    "Journal Theme (JAMA)"
  )
```

## {gtsummary} theme_gtsummary_language()

```{r}
#| output-location: "column"
reset_gtsummary_theme()
theme_gtsummary_language(language = "zh-tw")

trial |> 
  tbl_summary(
    by = trt, 
    include = c(age, response)
  ) |>
  add_p() |> 
  modify_caption(
    "Language Theme (Chinese)"
  )
```

**Language options:**

:::{.columns style="font-size:0.65em"}
::: {.column width="30%"}

- German 
- English
- Spanish 
- French 
- Gujarati
- Hindi

::: 

::: {.column width="30%"}
- Icelandic
- Japanese
- Korean
- Marathi
- Dutch

:::

::: {.column width="30%"}
- Norwegian
- Portuguese
- Swedish
- Chinese Simplified
- Chinese Traditional
:::
:::

## {gtsummary} theme_gtsummary_compact()

```{r}
reset_gtsummary_theme()
theme_gtsummary_compact()

trial |> 
  tbl_summary(
    by = trt, 
    include = c(age, response)
  ) |>
  modify_caption("Compact Theme")
```

*Reduces padding and font size*

## A pharma theme?

While not yet exported from gtsummary, we can create a theme for tables that look more like what we expect in pharma.

- Fixed-width font

- Continuous variable summaries default to multi-line

- Function for rounding percentages includes leading white space

- Default right alignment on summary statistics

```{r}
#| echo: false
theme_gtsummary_compact()
cards::ADSL |>
  dplyr::select(ARM, AGE, AGEGR1, SEX) |>
  dplyr::mutate(
    ARM = factor(ARM, levels = c("Placebo", "Xanomeline Low Dose", "Xanomeline High Dose")),
    # AGEGR1 = factor(AGEGR1, levels = c("<65", "65-80", ">80"), labels = c("&lt;65", "65-80", ">80"))
    AGEGR1 = factor(AGEGR1, levels = c("<65", "65-80", ">80"), labels = c("<65", "65-80", ">80"))
  ) |>
  tbl_summary(
    by = ARM,
    type = AGE ~ "continuous2",
    statistic = AGE ~ c("{median} ({p25}, {p75})", "{mean} ({sd})", "{min} - {max}"),
    value = SEX ~ "F",
    label = list(SEX = "Female", AGEGR1 = "Age Group"),
    digits = list(all_categorical() ~ list(p = cards::alias_as_fmt_fn("xx.x%")),
                  all_continuous() ~ 1L)
  ) |>
  add_stat_label() |>
  bold_labels() |>
  modify_header(all_stat_cols() ~ "**{level}**  \nN = {n}") |>
  modify_column_alignment(columns = -label, "right") |>
  as_gt() |>
  gt::cols_width(gtsummary::all_stat_cols() ~ gt::pct(23)) |>
  gt::opt_table_font(stack = "monospace-code")
```

## {gtsummary} set_gtsummary_theme()


* `set_gtsummary_theme()` to use a custom theme.

* See the {gtsummary} + themes vignette for examples

[http://www.danieldsjoberg.com/gtsummary/articles/themes.html](http://www.danieldsjoberg.com/gtsummary/articles/themes.html)

```{r}
#| echo: false
reset_gtsummary_theme()
```
